---
title: "Honeybee dominance reshapes floral resource use of wild pollinators across Europe"
format:
  pdf:
    colorlinks: true            # <-- turn on colored links
    linkcolor: RoyalBlue        # internal links (TOC, refs)
    citecolor: RoyalBlue        # in-text citations
    urlcolor: RoyalBlue         # URLs/DOIs
    include-in-header:
      text: |
        \usepackage[dvipsnames]{xcolor}  % provides RoyalBlue
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
        \usepackage{caption}
        \captionsetup[figure]{labelformat=empty}
        \captionsetup[table]{labelformat=empty}
        \usepackage{lineno}
        \linenumbers
bibliography: references.bib
link-citations: true            # note: YAML booleans are lowercase
latex-tinytex: false
links-as-notes: false
prefer-html: true
---


\vspace{-2cm}

\begin{center}

 \textbf{Lorenzo Marini\textsuperscript{1}} | \textbf{Jose B. Lanuza\textsuperscript{2}}
 
\end{center}


1 DAFNAE, University of Padova, Viale dell’Università 16, Legnaro, Padova, Italy

2 Mediterranean Institute for Advanced Studies (IMEDEA, UIB-CSIC), Mallorca, Balearic Islands, Spain


*E-mail:* [lorenzo.marini@unipd.it](mailto:lorenzo.marini@unipd.it), *tel.:* +39 049 827 2807

# ABSTRACT

*Apis mellifera* L. (the western honeybee) is one of the most abundant and frequent pollinator species occurring in plant-pollinator networks worldwide. Its high abundance is often associated with the introduction of managed colonies for honey production and crop pollination, raising concerns about potential impacts on wild pollinators due to resource competition and pathogen spill-over. Using a European-level plant-pollinator interaction dataset, we quantified resource overlap between honeybees and wild pollinators, and changes in diet specialisation along a gradient of honeybee dominance across Europe. Increasing dominance reduced resource use overlap between honeybees and wild pollinators indicating strong floral niche partitioning, yet wild pollinators became more generalist, visiting a greater diversity of floral resources. Using location-scale models, we also observed a more inconsistent and unpredictable behaviour across wild pollinators when honeybees were very abundant. Increasing local floral diversity provided foraging opportunities for wild pollinators, but it did not mitigate these effects. Our results suggest that avoidance and interference behaviours between managed and wild pollinators are common across Europe, but the importance of these resource use shifts for pollinator population dynamics, changes in community diversity, and ecosystem functioning in the long term remains an open question.

# INTRODUCTION

*Apis mellifera* L. (the western honeybee) is one of the most abundant and frequent pollinator species occurring in plant-pollinator networks worldwide across both agricultural and natural landscapes [@hung2018]. Its high abundance is often associated with the introduction of managed colonies for honey production and crop pollination [@gazzea2023; @visick2023], raising concerns about potential impacts on wild pollinators due to competition for floral resources [@iwasaki2022; @thomson2020] and pathogens’ spillover [@maurer2024]. Field studies manipulating colony density in the short term have shown that honeybees can often cause competitive displacement through the use of shared resources [@hudewenz2015; @magrach2025; @magrach2017; @pasquali2025; @valido2019], forcing other pollinators to forage on alternative floral resources that are less preferred by the honeybee [@page2023; @pasquali2025; @thomson2016; @thomson2004; @travis2025]. This finding is consistent with the low plant specialisation, opportunistic foraging behaviour, and high population density of this species [@lowe2023]. Manipulative experiments involving the exclusion or supplementation of honeybees are complex due to their widespread distribution and long foraging distances. As a result, we still lack a comprehensive synthesis to evaluate how common these processes are across different biogeographical regions, climates, and habitats.

Quantifying resource overlap between the honeybee and wild pollinators and their diet specialisation across ecological networks can provide a complementary approach to detect shifting preferences along a gradient of honeybee dominance [@cappellari2022]. When floral resources are limited and pollinators compete asymmetrically, low resource overlap and an altered foraging breadth should emerge, indicating potential competitive displacement in which a superior competitor forces inferior competitors to shift their niche and feed on sub-optimal floral resources [@magrach2017; @pasquali2025]. Such displacement can reduce the abundance of the inferior competitor [@lindstrom2016; @thomson2016], or, in the most extreme cases, lead to their local disappearance [@lazaro2021]. Other wild pollinators, such as bumblebees, can also displace the honeybee, which, under certain circumstances, becomes an inferior competitor [@balfour2015; @wignall2020]. Foraging shifts are expected to increase with honeybee density and floral resource availability in the network [@gomez2022]. Low plant diversity can increase resource overlap between honeybees and other pollinators due to the lack of alternative floral resources, forcing species to forage on the same plants [@bernhardsson2024; @page2024], while functional trait similarity (e.g., body size and tongue length) can also reshape resource use [@cappellari2022]. Insect pollinators are highly diverse, with Hymenoptera, Diptera, Lepidoptera and Coleoptera representing the key flower-visiting orders, which differ in life histories, nesting requirements, diet preferences, and floral resource specialisation. Yet most research has focused on the relationship between honeybees and wild bees [@iwasaki2022] with limited understanding of the responses of other key pollinator groups to increased honeybee dominance.  

Using the European plant-pollinator network database [@lanuza2025], this study provides the first continental synthesis of the effect of honeybee dominance on resource overlap and diet specialisation across more than 1800 species of wild pollinators, including wild bees, hoverflies, butterflies and beetles. Based on more than 649 independent networks and 1625 sampling events, the role of the honeybee across European plant-pollinator networks was assessed in terms of visitation rate patterns and floral resource use. It was tested whether increasing dominance of the honeybee in the networks broadened the foraging breadth of co-occurring wild pollinators and reduced their resource overlap with the honeybee, indicating a potential intensification of competition for floral resources. Using recently developed scale-location models, we also explored the variability in pollinator response. These relationships were examined across the major biogeographical, climatic, and land-use gradients in Europe. 

# MATERIALS AND METHODS

*Data*

This work uses the recently compiled database of European plant-pollinator networks [EuPPolNet\; @lanuza2025]. The harmonised taxonomy across the different networks enabled testing for resource overlap and diet specialisation at the species level across multiple networks. Weighted networks that met the following criteria were included: i) presence of honeybees, ii) more than one plant species, and iii) more than one pollinator species. The pollinator species were pooled into six non-overlapping taxonomic groups that are expected to interact differently with the honeybee due to distinct behavioural and morphological traits: wild bees (hymenopteran families: Apidae, Halictidae, Andrenidae, Colletidae, Megachilidae, Melittidae), non-bee hymenopterans (all families excluding wild bees), lepidopterans, coleopterans, hoverflies (dipteran family: Syrphidae), and non-syrphid flies (all dipteran families excluding Syrphidae). The final filtered dataset comprised 592 independent networks from 46 studies with a total of 1634 sampling events. 



*Resource overlap and diet specialisation*

For each sampling event per network, we used the observed interactions to compute the Morisita-Horn index [@horn1966] to quantify resource overlap between each focal pollinator species and the honeybee:


$$
\textit{Morisita-Horn} =
\frac{2 \sum_{k}^{S} p_k h_k}{
\left( \sum_{k}^{S} \left( \frac{p_k^2}{P^2} \right)
+ \sum_{k}^{S} \left( \frac{h_k^2}{H^2} \right) \right) {PH}
}
$$

Here, $p_k$ is the number of interactions of the focal pollinator species with each plant species k and $h_k$ is the number of interactions of the honeybee with each plant $k$. $S$ is the number of plant species, $P$ the total number of interactions of the focal pollinator, and $H$ is the total number of interactions of the honeybee. This symmetric index ranges from 0 (no shared use of resources) to 1 (identical resource use profile). The index normalises visits for each species by its total, using the relative distributions of resource use. Hence, it is not mathematically related to the abundance of focal pollinators or honeybees. The Müller index was not used as it is directional and mathematically related to honeybee dominance, ranging from 0 when honeybees are rare to 1 when they dominate the network, i.e. irrespective of the resource overlap, if honeybees are rare visitors overall, observed Müller index values will tend to 0, while if honeybees numerically dominate the network, Müller index values will inflate toward 1.  

For each pollinator, we also estimated two complementary measures of diet specialisation, normalised degree and PDI [@poisot2012]. Normalised degree is defined as $\frac{k_i}{N_{\text{plant}}}$, where k is the species degree (number of plant species visited by pollinator i) and Nplant is the total number of plant species in the network. As normalised degree is mathematically constrained by the number of plant species in a network, values are not directly comparable among networks with different plant species richness.

The pair difference index (PDI) was computed using its normalised version. For a species with interaction strengths $w_1 \ge w_2 \ge \dots \ge w_k$ across $k$ plants, we set $p_i = \frac{w_i}{\max(w)}$ so that $p_1 = 1$. The index is defined as:

$$
PDI_{\text{norm}} = \frac{\sum_{i}^{k} (1 - p_i)}{k - 1}
$$

The index lies between 0 and 1, with 0 indicating a perfect generalist (all links equal), and 1 a perfect specialist (only one link). The Kullback–Leibler distance [d'\; @bluthgen2006] was not used, as in the absence of independent data on plant abundance, the metric d' is sensitive to the behaviour of the other pollinator species in the network, and the metric becomes mathematically related to the abundance of any dominant species [@poisot2012].

*Null models*

Most node-level metrics computed across networks can be affected by overall network topology. Null models can be used effectively to compare metrics across networks, removing the effect of neutral processes [@bluthgen2024]. For each network and sampling event, we created 500 null models using the Patefield algorithm [@patefield1981]. This null model maintains the total interactions of both pollinator and plant species constant. We used this null model because the total number of interactions per pollinator reflects the local abundance of honeybees and other pollinators in the habitat, while keeping the total number of interactions per plant constant preserves the fact that some plant species are inherently more attractive to pollinators than others. While keeping these constraints, the null model randomly reshuffled pollinator visits across plants. For each species, mean and SD from the 500 null models were computed for all metrics and used to compute standardised effect sizes: $SES = \frac{(\text{observed} - \text{mean}_{\text{null}})}{SD_{\text{null}}}$. Due to the generating constraints of the Patefield algorithm, in some networks, SESs could not be computed because SD approximated 0. In further analyses, networks with SD = 0 or IQR = 0 were removed as SES could not be defined. 


*Drivers*

In the EuPPollNet database, each network contains information on geographical coordinates, sampled habitat, date of the sampling, and biogeographical regions. Summer mean temperature was derived by averaging the monthly averages from May to September for each sampling site (i.e., network) using the Chelsea database [@karger2023]. Since there was a strong collinearity between climate and habitats, the original habitat types were grouped in two broad categories: (i) natural habitats including forest/woodland, riparian vegetation, semi-natural grasslands, moors and heathland, sclerophyllous vegetation, montane to alpine grasslands, beaches, dunes, sands, riparian; and (ii) anthropogenic habitats including intensive grasslands, agricultural land, agricultural margins, green urban areas, and ruderal vegetation. At the network level and for each sampling event, we computed the total number of plant partners, the total number of pollinator species, and the honeybee dominance as the ln-ratio of total honeybee interactions and non-honeybee interactions. The ln-ratio is 0 when honeybee interactions are equal to the sum of the interactions of all other pollinators sampled. We could not use the total abundance of honeybees because sampling effort differed across studies. Although the database includes information on floral abundance, inconsistent sampling methodologies limited its use, and only 74% of the studies reported such data.

**Data analysis**

*Resource overlap with the honeybee and foraging breadth of wild pollinators*

The three metrics (i.e., Morisita-Horn index, normalised degree, and PDI) were transformed into standardised effect sizes (SESs) and used as response variables in generalised linear mixed models (GLMMs) with a Gaussian distribution in the glmmTMB package [@mcgillycuddy2025]. For each response, a location-scale model was fitted. Although ecological data often exhibit non-constant variance, this pattern is commonly treated as noise that violates the assumptions of statistical models, while heteroscedasticity can sometimes reflect meaningful ecological processes [@cleasby2011]. The location component (conditional in glmmTMB) tested the effect of pollinator group, honeybee dominance, number of plant partners, temperature and habitat type on the SES mean of each metric. A quadratic term for dominance was also tested, and the term was removed when linearity could not be rejected (P > 0.05). Two interactions were also tested: honeybee dominance with pollinator taxonomic groups and honeybee dominance with floral richness. The first tested whether the six pollinator taxonomic groups responded differently to the honeybees, while the second tested whether the availability of floral partners modified the honeybee dominance effect on wild pollinators. The sampling date nested within network ID, nested within Study ID, was used as a random structure. Second, a scale component (dispformula in glmmTMB) was used to test whether the number of floral partners and honeybee dominance affected the variance of the SES residuals. The same random structure was included in the scale component, but it was fitted independently from the conditional, as correlated random effects are not implemented in glmmTMB. Models with and without the scale component were compared with the likelihood ratio test using AIC. In all models, LRTs always supported the inclusion of the two predictors and the random structure in the scale component. We tested for potential collinearity among predictors using a global model with no interactions using the check_collinearity() function in the performance package [@ludecke2021]. All predictors included in the modelling exhibited low VIFs  (< 1.5). Continuous predictors were scaled using the scale() function. Model diagnostics were performed using the DHARMa package [@hartig2016]. All models performed well except for Morisita-Horn, which presented a slight deviation (Supplementary Information). All analyses were run using R version 4.5.1. 

*Sensitivity analysis*

As resource overlap and diet specialisation metrics could be influenced by species degree and abundance, a sensitivity analysis was conducted, where focal pollinator species were removed based on their total number of individuals per network and sampling event. Main text analyses included only pollinator species with at least five individuals and a minimum of five honeybees per network. In the Supplementary Information, we presented the same models with different thresholds of minimum number of individuals per network per sampling event (min = 2, 3, 5, and 10). Results were qualitatively similar and showed strong consistency in the observed patterns.

# RESULTS

*General results*

Honeybees visited approximately 50% of all plant species in the dataset. Only 3.5% of plants were visited exclusively by the honeybee. They contributed on average 36% of the interactions per network (**Fig 1 and Figs. S1-S2**), ranging from networks where they were present only sporadically and at low density (under 1% of visits) to networks where they were consistently the main visitor (up to 98% of visits). As expected, there was a positive linear association between floral richness and pollinator richness across networks (**Fig. 2**). Both floral and pollinator species richness showed a weak correlation with honeybee normalised degree, with a trend for higher normalised degree at low species richness values. In the sampled networks, honeybee dominance was weakly negatively correlated with floral and pollinator richness and positively with honeybee normalised degree.


```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="85%", fig.height=6.5, fig.width=6.5, fig.cap= "\\textbf{Figure 1}. Approximate locations of the studies, showing the proportion of \\textit{Apis} interactions weighted by the total number of pollinator species per study."}

library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(giscoR)
library(ggtext)


#--------------------------------------------------
# 0. Load data
#--------------------------------------------------

df     <- readRDS("Data/filtered_interaction_data.rds")
lanuza <- readRDS("Data/raw_interaction_data.rds")

#--------------------------------------------------
# 1. Get Study IDs and coordinates
#--------------------------------------------------

id <- df %>%
  separate(id, into = c("Study_id", "Location", "Date"), sep = "//") %>%
  distinct(Study_id) %>%
  pull()

coordinates_sf <- lanuza %>%
  filter(Study_id %in% id) %>%
  select(Study_id, Latitude, Longitude) %>%
  distinct() %>%
  group_by(Study_id) %>%
  slice(1) %>%
  ungroup() %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  st_transform(3035)

#--------------------------------------------------
# 2. Pollinator richness (number of pollinators per study)
#--------------------------------------------------

pollinator_richness <- df %>%
  separate(id, into = c("Study_id", "Location", "Date"), sep = "//") %>%
  group_by(Study_id) %>%
  summarise(
    n_pollinators = n_distinct(pollinator),
    .groups = "drop"
  )

#--------------------------------------------------
# 3. Study-level Apis vs Other interactions
#--------------------------------------------------

apis_wide <- df %>%
  separate(id, into = c("Study_id", "Location", "Date"), sep = "//") %>%
  mutate(
    is_apis = if_else(
      grepl("Apis", pollinator, ignore.case = TRUE),
      "Apis",
      "Other"
    )
  ) %>%
  group_by(Study_id, is_apis) %>%
  summarise(
    n_interactions = sum(interaction, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from  = is_apis,
    values_from = n_interactions,
    values_fill = 0
  )

#--------------------------------------------------
# 4. Combine coordinates, interactions, and richness
#--------------------------------------------------

pie_data <- coordinates_sf %>%
  left_join(apis_wide, by = "Study_id") %>%
  left_join(pollinator_richness, by = "Study_id") %>%
  mutate(
    x = st_coordinates(geometry)[, 1],
    y = st_coordinates(geometry)[, 2],
    total_inter     = Apis + Other,
    prop_apis_study = if_else(total_inter > 0, Apis / total_inter, NA_real_)
  ) %>%
  st_drop_geometry()

# Convert back to sf for spatial operations
pie_sf <- st_as_sf(
  pie_data,
  coords = c("x", "y"),
  crs    = 3035,
  remove = FALSE
)

#--------------------------------------------------
# 5. Countries layer (for reference only)
#--------------------------------------------------

cntries <- gisco_get_countries(region = "Europe", year = "2016") %>%
  st_transform(3035)

#--------------------------------------------------
# 6. Build FULL rectangular grid over Europe extent
#    (NOT clipped by borders)
#--------------------------------------------------

bbox_grid <- st_bbox(c(
  xmin = 2200000,
  xmax = 7150000,
  ymin = 1380000,
  ymax = 5500000
), crs = st_crs(3035))

grid_poly <- st_make_grid(
  st_as_sfc(bbox_grid),
  cellsize = 100000,   # 100 km grid
  square   = TRUE
)

grid_sf <- st_sf(
  grid_id  = seq_along(grid_poly),
  geometry = grid_poly
)

#--------------------------------------------------
# 7. Assign studies to grid cells & aggregate
#    (weight by number of pollinators per study)
#--------------------------------------------------

# Join each point to the grid cell it falls in
join <- st_join(pie_sf, grid_sf, join = st_within)

# Weighted aggregation per grid cell
grid_agg <- join %>%
  st_drop_geometry() %>%
  group_by(grid_id) %>%
  summarise(
    prop_apis = if (all(is.na(prop_apis_study)) || all(is.na(n_pollinators))) {
      NA_real_
    } else {
      weighted.mean(prop_apis_study, w = n_pollinators, na.rm = TRUE)
    },
    n_studies    = n(),
    total_weight = sum(n_pollinators, na.rm = TRUE),
    .groups = "drop"
  )

# Attach aggregated values back to all grid cells
grid_full <- grid_sf %>%
  left_join(grid_agg, by = "grid_id") %>%
  mutate(
    prop_apis_perc = prop_apis * 100   # convert to percentage for plotting
  )

#--------------------------------------------------
# 8. Get centroids for symbol plotting
#--------------------------------------------------

grid_centroids <- st_centroid(grid_full)

grid_centroids_df <- grid_centroids %>%
  mutate(
    cx = st_coordinates(geometry)[, 1],
    cy = st_coordinates(geometry)[, 2]
  ) %>%
  st_drop_geometry()

#--------------------------------------------------
# 9. Plot: background grid + size-varying squares
#--------------------------------------------------
# 8b. Breaks for legends

size_limits <- range(grid_centroids_df$total_weight, na.rm = TRUE)
size_breaks <- pretty(size_limits, n = 4)

ggplot() +
  # Country polygons as reference
  geom_sf(
    data  = cntries,
    fill  = "grey85",
    color = "black",
    size  = 0.1
  ) +
  # Light background grid (all same size, low alpha)
  geom_sf(
    data  = grid_full,
    fill  = NA,
    color = "grey90",
    size  = 0.1
  ) +
  # Squares at centroids: size = total pollinator richness per cell,
  # fill = weighted proportion of Apis interactions
  geom_point(
    data = grid_centroids_df,
    aes(
      x    = cx,
      y    = cy,
      size = total_weight,
      fill = prop_apis_perc
    ),
    shape  = 22,  # filled square
    color  = "black",
    stroke = 0.1
  ) +
  coord_sf(
    xlim   = c(2200000, 7150000),
    ylim   = c(1380000, 5500000),
    expand = FALSE
  ) +
  scale_fill_viridis_c(
    option   = "C",
    na.value = NA,
    name     = "Proportion of *Apis*\ninteractions (%)",
    breaks   = c(0,20,40,60),
    labels   = c(0,20,40,60),
    limits   = c(0, 70), 
  ) +
  scale_size_continuous(
    name = "Total pollinator\nrichness",
    range = c(1, 6.5)  # tweak as needed
  ) +
  guides(
    fill = guide_colorbar(order = 1),
    size = guide_legend(order = 2)
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.background = element_rect(fill = NA),
    panel.border     = element_rect(colour = "black", fill = NA, size = 1),
    legend.position  = "right",
    legend.key.size = unit(0.5, "lines"),
    axis.ticks       = element_line(color = "black"),
  legend.title    = element_markdown(size = 10)
  ) +
  xlab(NULL) +
  ylab(NULL)


```


*Resource overlap*

The standardised effect sizes (SESs) of Morisita-Horn were mainly negative, indicating that in most cases, wild pollinators co-occurring with the honeybee exhibited a lower resource overlap than expected from the null models. The location component on the mean SES supported an interaction between pollinator group and honey bee dominance (**Table S1a**), with a steeper slope for coleopterans and syrphids compared to other groups (**Fig. 2a**). The interaction between the number of plants and dominance was not strongly supported, while the number of plants available in the network reduced resource overlap. No effect of temperature or habitat was found. The scale component indicated an increase in the SES residual variance with dominance and a decrease with floral richness (**Table S1b**), meaning that pollinator resource overlap became more variable at high dominance and low floral richness.


```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="85%", fig.height = 5.5, fig.width = 5.5,  fig.cap= "\\textbf{Figure 2}. Scatterplots showing the relationships between the two components of network size (i.e., plant and pollinator species richness) and honeybee normalized degree and dominance (log ratio of honeybee interactions over total interactions). The dashed orange line represents the linear regression, and the solid black line represents the GAM fitted with cubic regression splines. Each point corresponds to an independent sampling event within a network. Only networks with at least five honeybee individuals were included. Diagonal panels represent variable density distributions, while upper panels indicate Pearson correlation coefficients between metrics."}

library(GGally)
library(ggplot2)
library(dplyr)

#### Load data ####
apis_network_metrics = readRDS("Data/apis_network_metrics.rds")


######## FIG. 1 COVARIATION BETWEEN DOMINANCE, ND, N POLL AND N PLANT ####

apis_summary2_no1 = apis_network_metrics%>%
  filter(Apis_interactions>1)

vars = c("plant_n","pollinator_n","Apis_ndegree", "Dominance")
df_subset = apis_summary2_no1[, vars, drop = FALSE]

theme_set(
  theme_bw(base_size = 9) +
    theme(
      panel.grid.minor = element_blank(),
      # panel.grid.major = element_blank(),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
      strip.text = element_text(face = "bold", margin = margin(t = 4, r = 4, b = 4, l = 4)),
      strip.background = element_rect(
        fill = "gray80",
        color = "white",
        linewidth = 0.9,
        size=1),
      
      panel.spacing = unit(0.6, "lines")))
# Custom lower panel: GAM + linear
lower_gam_lm = function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_point(shape = 16, color = "lightblue", size = 0.8, alpha = 0.9) +  # shape=16 removes border
    geom_smooth(method = "lm", se = FALSE, color = "#FA9E3BFF", linetype = "dashed", ...) +
    geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = FALSE, color = "black", ...) +
    scale_x_continuous(n.breaks = 4) +   # limit x-axis ticks
    scale_y_continuous(n.breaks = 4) 

}

p = ggpairs(
  df_subset,
  columns = c("plant_n","pollinator_n","Apis_ndegree",  "Dominance"),
  columnLabels = c("Plant richness", "Pollinator richness", "Hb normalized degree",  "Hb dominance"),
  lower = list(continuous = wrap(lower_gam_lm)),
  diag = list(continuous = wrap("densityDiag", alpha = 0.5, fill = "lightblue")),
  upper = list(continuous = wrap("cor", size = 2.5))
  #diag = list(continuous = wrap(custom_density_diag))
) 

p 
```

*Diet specialization*

The SESs for the paired difference index (PDI) were mostly positive or near 0, indicating that pollinator specialisation tended to be relatively higher than or similar to the expected values from the null models. The location-scale component indicated that mean PDI declined non-linearly with increasing honeybee dominance for all groups, except for the coleopterans, which showed an opposite pattern (**Fig. 2B**). Coleopterans and lepidopterans were those with higher specialisation. A weak interaction between dominance and floral richness on mean PDI, with a slightly steeper negative slope in networks with more flowering species. Networks sampled in warmer regions presented higher pollinator specialisation (**Table S2a**). The dispersion component indicated a negative effect of floral richness on the PDI residual variability, while dominance did not have an effect (**Table S2b**). 

The SESs of normalised degree were mainly negative or near 0, indicating that the pollinator use of the available plants tended to be relatively lower than or similar to the expected values from the null models. Consistent with PDI, an interaction between taxonomic group and dominance was found (**Table S3a**), with normalised degree increasing with honeybee dominance in all groups except for coleopterans, which exhibited an opposite pattern (**Fig. 2C**). An interaction between floral richness and dominance was also observed, where mean normalised degree declined with increasing floral richness but only at low dominance values. The dispersion component indicated an increase in the SES residual variance of normalised degree with dominance and a decrease with floral richness (**Table S3b**).

*Model performance and sensitivity*

The residuals’ diagnostics for the four scale-location models were satisfactory, given the complexity of the data and the modelling. The deviations were minor, but statistically significant due to the large number of observations. The sensitivity analysis indicated that removing species with different thresholds of minimum interactions per network (n = 2, and 10) did not alter the observed patterns (**Tables S4-S5-S6** and **Figure S2** for a threshold of 2 interactions, and **Tables S7-S8-S9** and **Figure S3** for a threshold of 10 interactions). Results of these sensitivity analyses are reported in Supplementary materials.

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%", fig.height = 2.5, fig.width = 8,  fig.cap= "\\textbf{Figure 3}. Effect of honeybee dominance on the standardised effect size (SES) of A) resource overlap with the honeybee (Morisita-Horn index), B) Paired difference index (PDI) and C) normalised degree of wild pollinators across different groups of flower-visiting insects. Lines represent predicted values from location-scale glmm. The analysis included networks with at least 5 honeybee individuals and wild pollinator species with at least 5 individuals per sampling event."}

library(ggeffects)
library(ggplot2)
library(patchwork)
library(glmmTMB)

# Load models
fit_SES_pdi <- readRDS("Data/fit_SES_pdi.rds")
fit_SES_nd  <- readRDS("Data/fit_SES_nd.rds")
fit_SES_mh  <- readRDS("Data/fit_SES_mh.rds")

# Base theme
base_theme <- theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.key            = element_rect(fill = "white", colour = NA),
    legend.background     = element_blank(),
    legend.box.background = element_blank()
  )

#my_cols <- viridisLite::viridis(6)
#my_cols <-scico::scico(6, palette = "roma")
#my_cols <-scico::scico(6, palette = "hawaii")
my_cols <-  c(
  "#000000",  # black
  "#E69F00",  # orange
  #"#56B4E9",  # sky blue
  "#009E73",  # bluish green
  "#F0E442",  # yellow
  "#0072B2",  # blue
#  "#D55E00",  # vermillion
  "#CC79A7"   # reddish purple
)



# Marginal effects
gp_mh  <- ggpredict(fit_SES_mh,  terms = c("Dominance[all]", "Group"), type = "fixed")
gp_pdi <- ggpredict(fit_SES_pdi, terms = c("Dominance[all]", "Group"), type = "fixed")
gp_nd  <- ggpredict(fit_SES_nd,  terms = c("Dominance[all]", "Group"), type = "fixed")

# Plots with different linetypes per group
mh <- plot(gp_mh) +
  aes(linetype = group) +
    scale_color_manual(values = my_cols) +   
  labs(
    x = "",
    y = "SES Morisita-Horn",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",  # don't show CI fill as a separate legend
    colour   = guide_legend(override.aes = list(fill = NA)),   # <-- remove grey box
    linetype = guide_legend(override.aes = list(fill = NA))    # <-- same for linetype legend
  ) +
  ggtitle("")

pdi <- plot(gp_pdi) +
  aes(linetype = group) +
  scale_color_manual(values = my_cols) +   
  labs(
    x = "Honeybee dominance (ln-ratio)",
    y = "SES PDI",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",
    colour   = guide_legend(override.aes = list(fill = NA)),
    linetype = guide_legend(override.aes = list(fill = NA))
  ) +
  ggtitle("")

nd <- plot(gp_nd) +
  aes(linetype = group) +
  scale_color_manual(values = my_cols) +   
  labs(
    x = "",
    y = "SES normalised degree",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",
    colour   = guide_legend(override.aes = list(fill = NA)),
    linetype = guide_legend(override.aes = list(fill = NA))
  ) +
  ggtitle("")

# Combine and apply theme
p <- (mh | pdi | nd) +
  plot_layout(ncol = 3, nrow = 1, guides = "collect") &
  base_theme &
  theme(
    legend.position = "right",
    base_size = 7.5,
    panel.spacing = unit(0.1, "lines"),
    plot.margin = unit(c(0, 0, 0, 0), "pt"),
    plot.tag = element_text(face = "bold", vjust = 0.35),
    legend.key.width = unit(1.4, "cm")
  )

p + plot_annotation(tag_levels = "A", tag_suffix = ")")

```

# DISCUSSION

Honeybees were dominant visitors across European plant-pollinator networks, and their increasing dominance was associated with a systematic decline in floral resource overlap with co-occurring wild pollinators, as well as a broader and more variable foraging breadth among co-occurring species. Although the observed associations were consistent across natural and disturbed habitats and climatic gradients, the major pollinator groups did not respond equally to honeybee dominance. As honeybees are known to remove a large share of floral resources, often from the most abundant and rewarding flowering species [@hung2018], this finding suggests that honeybees can reshape the resource use of co-occurring wild pollinator species [@magrach2025]. How important these diet shifts are for population dynamics, in community diversity, or ecosystem functioning remains an open question [@beaurepaire2025; @pascual2024].

The observed low interactions of wild pollinators with the flowering species visited by the honeybee may reflect a direct response to resource depletion, but possibly also to deterrent scent cues or other indicators of honeybee presence [@stout2001]. Although the direction and the mechanisms underlying this avoidance behaviour are still not fully understood, reduced resource use overlap between the honeybee and other wild bees has been reported in many experimental and observational studies [@cappellari2022; @page2023; @magrach2025]. Manipulative studies that have excluded honeybees have demonstrated that a reduced resource overlap is often associated with short-term foraging shifts that can be reversed when honeybees are removed from the community [@pasquali2025]. This pattern might also be affected by plant diversity, where increasing plant niche space generally reduces the probability of intra-species contact, providing more foraging opportunities to co-occurring species [@gomez2022]. The inclusion of a large number of networks with a harmonised taxonomy allowed us to test whether this reduced overlap changed for the major European pollinator groups. On the one hand, wild bees and lepidopterans exhibited a relatively higher overlap with the honeybee than other groups, consistent with the similar nectar requirements of these taxa. On the other hand, coleopterans exhibited the lowest overlap. Observed coleopterans, such as Scarabaeidae (Cetoniinae), Oedemeridae, Nitidulidae, and Melyridae, are often florivores that, besides removing pollen and nectar, frequently damage flower parts, thereby reducing flower attractiveness for pollinators [@haas2020]. Coleopterans usually select flower species mainly based on their greater longevity, spending a long time feeding on the same species [@camurca2024]. Hence, it is likely that the observed reduced overlap was a consequence of honeybees avoiding flowers either damaged or occupied by florivores [@boaventura2022]. In addition, we cannot exclude the possibility that other species can also push the honeybee away from preferred resources [@inouye1978; @taggar2021]. 


The scale-location model also indicated that not only the mean resource overlap was affected by the honeybee, but also its variability. This contrasts with many ecological and behavioural studies where variability is often assumed to be constant, potentially masking informative environmental and biological patterns in the data [@nakagawa2025]. An increase in behavioural variability in pollinators in the presence of a dominant pollinator is consistent with a trait-mediated shift in behaviour, i.e. species differed in their ability to avoid honeybee-used plants (e.g. tongue length, body size, handling time, floral preferences) and under intense honeybee pressure, some species shifted away (low overlap), while others could not avoid or even track the same rewarding resources due to their trait similarity with the honeybee [@cappellari2022].

Optimal foraging theory also suggests that when resources are abundant and diverse and species are not competing strongly, pollinators should concentrate on a few most-preferred plants and have narrow food niches, while increased competition usually forces inferior competitors to become more generalist in their resource use [@macarthur1966]. After removing the effects of neutral processes using null models, we observed a general decrease in diet specialisation with increasing honeybee dominance for all pollinator taxonomic groups, except for florivorous coleopterans. Again, pollinator taxonomic groups showed contrasting responses likely due to trait-mediated differences in their ability to adjust their foraging behaviour under increased honeybee dominance. An interaction between floral diversity and honeybee dominance was also found, with a more pronounced specialisation with increasing floral diversity only at low dominance, while at high dominance levels, specialisation did not depend on floral diversity. Again, the location-scale model also detected an increase in the variability of responses among species, suggesting that honeybee pressures might amplify intra- and interspecific differences, likely driven by morphological and behavioural traits affecting resource use [@cappellari2022].

Although species can adjust to competition by shifting their diets, the coexistence of a superabundant pollinator such as the honeybee and native pollinators can be fragile and depend on floral resource availability only in part. It is important to stress that asymmetric competition among pollinators is probably the rule across plant-pollinator networks [@sponsler2021], and it has been observed in many other pollinator species besides the honeybee [@inouye1978; @taggar2021]. However, several studies examining nectar and pollen landscapes have indicated that honeybee-induced diet shifts are often associated with poorer nutritional quality in wild pollinator diets [@page2024; @wignall2020]. Hence, the effectiveness of conservation interventions for wild pollinators, in particular across agricultural landscapes, should always consider the potential competition with the honeybees [@bommarco2021; @raderschall2022]. At the same time, maintaining honeybee populations across agricultural landscapes is crucial for delivering adequate crop pollination. Under these circumstances, it is vital to identify the minimum number of honeybee hives to balance yield and biodiversity by providing needed pollination for the crop while minimising competition with wild pollinators [@fijen2022]. From a wild plant pollination perspective, the consequences of these diet shifts are difficult to predict and are expected to be strongly context-dependent [@gao2025; @magrach2017; @pascual2024; @valido2019]. Ultimately, the observed avoidance behaviours and the dominant role of honeybees suggest that interactions within pollinator communities are far from neutral and that more research is needed to understand the direction and mechanisms of potential antagonistic interactions. Although managed honey bees per se may not be a conservation priority, beekeeping is deeply rooted in the human-nature history [@sponsler2021] and is associated with key cultural services, highlighting the need to co-create effective, context-relevant strategies to reconcile beekeeping and the conservation of wild pollinators. 

# ACKNOWLEDGEMENTS

The research was supported by the European Union’s Horizon 2020 329 project “Safeguard” (Grant agreement ID: 101003476) funded under Societal Challenges - Climate 330 action, Environment, Resource Efficiency and Raw Materials to L.M.

# AUTHORS' CONTRIBUTIONS

Conceptualization: L.M., Formal analysis: L.M. & J.B.L., Resources: J.B.L., Writing - Original Draft: L.M., Writing - Review & Editing: L.M. & J.B.L., Supervision: L.M., Funding Acquisition: L.M.

# REFERENCES

::: {#refs}
:::

\newpage

# SUPPLEMENTARY INFORMATION

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the MH table (Conditional + Dispersion)
df_mh <- dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_mh, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_mh, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_mh <- df_mh %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_mh, is.numeric)
df_mh[num_cols] <- lapply(df_mh[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_mh$Component == "Conditional")
idx_disp <- which(df_mh$Component == "Dispersion")

df_mh = df_mh %>% select(!Component)

# Render table
kbl(
  df_mh,
  caption = "\\textbf{Table S1}. Scale-location model testing multiple predictors on the standardized effect size of pairwise resource overlap (Morisita–Horn) between the honeybee and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
   column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_mh), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))
```

\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the PDI table (Conditional + Dispersion)
df_pdi = dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_pdi, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_pdi, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_pdi = df_pdi %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_pdi, is.numeric)
df_pdi[num_cols] <- lapply(df_pdi[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_pdi$Component == "Conditional")
idx_disp <- which(df_pdi$Component == "Dispersion")

df_pdi = df_pdi %>% select(!Component)

# Render table
kbl(
  df_pdi,
  caption = "\\textbf{Table S2}. Scale-location model testing multiple predictors on the standardized effect size of paired difference index (PDI) between the honeybee and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
  column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_pdi), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))

```

\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the ND table (Conditional + Dispersion)
df_nd = dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_nd, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_nd, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_nd = df_nd %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_nd, is.numeric)
df_nd[num_cols] <- lapply(df_nd[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_nd$Component == "Conditional")
idx_disp <- which(df_nd$Component == "Dispersion")

df_nd = df_nd %>% select(!Component)

# Render table
kbl(
  df_nd,
  caption = "\\textbf{Table S3}. Scale-location model testing multiple predictors on the standardized effect size of normalised degree between the honeybee and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
  column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_nd), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))

```

\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

#overwrite data with sensitivity analysis
fit_SES_mh  = readRDS("Data/fit_SES_mh_over1.rds")

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the MH table (Conditional + Dispersion)
df_mh <- dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_mh, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_mh, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_mh <- df_mh %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_mh, is.numeric)
df_mh[num_cols] <- lapply(df_mh[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_mh$Component == "Conditional")
idx_disp <- which(df_mh$Component == "Dispersion")

df_mh = df_mh %>% select(!Component)

# Render table
kbl(
  df_mh,
  caption = "\\textbf{Table S4}. Scale-location model from the sensitivity analysis applying a minimum interaction threshold of two species. The model tests multiple predictors on the standardized effect size of pairwise resource overlap (Morisita–Horn) between honeybees and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
   column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_mh), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))
```


\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

#overwrite data with sensitivity analysis
fit_SES_pdi  = readRDS("Data/fit_SES_pdi_over1.rds")

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the PDI table (Conditional + Dispersion)
df_pdi = dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_pdi, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_pdi, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_pdi = df_pdi %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_pdi, is.numeric)
df_pdi[num_cols] <- lapply(df_pdi[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_pdi$Component == "Conditional")
idx_disp <- which(df_pdi$Component == "Dispersion")

df_pdi = df_pdi %>% select(!Component)

# Render table
kbl(
  df_pdi,
  caption = "\\textbf{Table S5}. Scale-location model from the sensitivity analysis applying a minimum interaction threshold of two species. The model tests multiple predictors on the standardized effect size of paired difference index (PDI) between honeybees and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
  column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_pdi), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))

```

\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

#overwrite data with sensitivity analysis
fit_SES_nd  = readRDS("Data/fit_SES_nd_over1.rds")

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the ND table (Conditional + Dispersion)
df_nd = dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_nd, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_nd, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_nd = df_nd %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_nd, is.numeric)
df_nd[num_cols] <- lapply(df_nd[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_nd$Component == "Conditional")
idx_disp <- which(df_nd$Component == "Dispersion")

df_nd = df_nd %>% select(!Component)

# Render table
kbl(
  df_nd,
  caption = "\\textbf{Table S6}. Scale-location model from the sensitivity analysis applying a minimum interaction threshold of two species. The model tests multiple predictors on the standardized effect size of normalised degree between honeybees and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
  column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_nd), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))

```

\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

#overwrite data with sensitivity analysis
fit_SES_mh  = readRDS("Data/fit_SES_mh_over9.rds")

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the MH table (Conditional + Dispersion)
df_mh <- dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_mh, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_mh, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_mh <- df_mh %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_mh, is.numeric)
df_mh[num_cols] <- lapply(df_mh[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_mh$Component == "Conditional")
idx_disp <- which(df_mh$Component == "Dispersion")

df_mh = df_mh %>% select(!Component)

# Render table
kbl(
  df_mh,
  caption = "\\textbf{Table S7}. Scale-location model from the sensitivity analysis applying a minimum interaction threshold of nine species. The model tests multiple predictors on the standardized effect size of pairwise resource overlap (Morisita–Horn) between honeybees and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
   column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_mh), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))
```


\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

#overwrite data with sensitivity analysis
fit_SES_pdi  = readRDS("Data/fit_SES_pdi_over9.rds")

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the PDI table (Conditional + Dispersion)
df_pdi = dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_pdi, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_pdi, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_pdi = df_pdi %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_pdi, is.numeric)
df_pdi[num_cols] <- lapply(df_pdi[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_pdi$Component == "Conditional")
idx_disp <- which(df_pdi$Component == "Dispersion")

df_pdi = df_pdi %>% select(!Component)

# Render table
kbl(
  df_pdi,
  caption = "\\textbf{Table S8}. Scale-location model from the sensitivity analysis applying a minimum interaction threshold of nine species. The model tests multiple predictors on the standardized effect size of paired difference index (PDI) between honeybees and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
  column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_pdi), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))

```

\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align='center', fig.height=1, fig.width=10}

#### Load libs ####
library(glmmTMB)     # make sure this is attached
library(dplyr)
library(knitr)
library(kableExtra)

#overwrite data with sensitivity analysis
fit_SES_nd  = readRDS("Data/fit_SES_nd_over9.rds")

# Helper that never touches the model frame
tab_df_glmmTMB <- function(fit, component = c("cond","disp"), label = NULL) {
  component <- match.arg(component)
  sm <- summary(fit)$coefficients[[component]]
  stopifnot(!is.null(sm))  # fail clearly if the component doesn't exist
  tibble::tibble(
    Term      = rownames(sm),
    Estimate  = sm[, "Estimate"],
    SE        = sm[, "Std. Error"],
    z         = sm[, "z value"],
    p         = sm[, "Pr(>|z|)"],
    Component = if (is.null(label)) component else label
  )
}

# Build the ND table (Conditional + Dispersion)
df_nd = dplyr::bind_rows(
  tab_df_glmmTMB(fit_SES_nd, component = "cond", label = "Conditional"),
  tab_df_glmmTMB(fit_SES_nd, component = "disp", label = "Dispersion")
)

# (Optional) clean prefixes if Term contains "Conditional_" or "Dispersion_"
df_nd = df_nd %>%
  mutate(
    Term = gsub("(?i)^(conditional|dispersion)[_ :.-]+", "", Term, perl = TRUE),
    Component = factor(Component, levels = c("Conditional", "Dispersion"))
  ) %>%
  arrange(Component)

# Round numeric columns
num_cols <- sapply(df_nd, is.numeric)
df_nd[num_cols] <- lapply(df_nd[num_cols], function(x) round(x, 3))

# Indices for grouping rows in kableExtra
idx_cond <- which(df_nd$Component == "Conditional")
idx_disp <- which(df_nd$Component == "Dispersion")

df_nd = df_nd %>% select(!Component)

# Render table
kbl(
  df_nd,
  caption = "\\textbf{Table S9}. Scale-location model from the sensitivity analysis applying a minimum interaction threshold of nine species. The model tests multiple predictors on the standardized effect size of normalised degree between honeybees and wild pollinator species across European networks. The conditional component tested the effects of group, honeybee dominance (linear and quadratic terms), temperature, habitat, plant richness, and the interactions between dominance and group and between dominance and plant richness on mean PDI. The dispersion component tested the effects of dominance and plant richness on the residual variance. Both components included pollinator species as a random intercept. Each observation corresponded to a species with at least five individuals recorded in a sampling event that also contained at least five honeybees.",
  booktabs = TRUE,
  escape = TRUE
) %>%
  column_spec(1, width = "18em") %>%       # Term column
  column_spec(2:ncol(df_nd), width = "4em") %>%  # numeric columns
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("(a) Conditional (mean)", min(idx_cond), max(idx_cond)) %>%
  pack_rows("(b) Dispersion (variance)", min(idx_disp), max(idx_disp))

```


\newpage

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="80%", fig.cap= "\\textbf{Figure S1}. Histogram showing the mean proportion of interactions of honeybees per network. The average value across all networks is indicated with a vertical dashed red bar."}


library(dplyr)
library(ggplot2)
filtered_interaction_data = readRDS("Data/filtered_interaction_data_all_cols.rds")


hb_percent = filtered_interaction_data %>% 
  mutate(Study_network_id = paste0(Study_id, Network_id)) %>% 
  mutate(group = if_else(grepl("Apis", pollinator, ignore.case = TRUE), "Honeybee", "Other")) %>% 
  group_by(Study_network_id, group) %>% 
  summarise(
    n_interactions = sum(interaction, na.rm = TRUE),   # sum of interactions per group
    .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from = group,
    values_from = n_interactions,
    values_fill = 0) %>%
  mutate(
    total = Honeybee + Other,
    perc_honeybee = 100 * Honeybee / total)

#Percentage of plants visited by the HB
# Calculate the mean
mean_hb <- mean(hb_percent$perc_honeybee, na.rm = TRUE)

theme_set(
  theme_bw(base_size = 9) +
    theme(
      panel.grid.minor = element_blank(),
      # panel.grid.major = element_blank(),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
      panel.spacing = unit(0.6, "lines")))

# Plot histogram + mean line
ggplot(hb_percent, aes(x = perc_honeybee)) +
  geom_histogram(
    bins = 30,                     # adjust bin number if needed
    fill = "skyblue", 
    color = "white"
  ) +
  geom_vline(
    xintercept = mean_hb,
    color = "red",
    linetype = "dashed",
    linewidth = 1
  ) +
  annotate(
    "text",
    x = mean_hb,
    y = 1,
    label = paste0("Mean = ", round(mean_hb, 1), "%"),
    vjust = -0.5,
    hjust = -0.1,
    color = "black",
    size = 4
  ) +
  labs(
    x = "Percentage of honeybee interactions per network",
    y = "Number of networks") 

```

\newpage

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%", fig.height = 2.5, fig.width = 8,  fig.cap= "\\textbf{Figure S2}. Effect of honeybee dominance on the standardised effect size (SES) of A) resource overlap (Morisita-Horn), B) PDI, and C) normalised degree, shown for the sensitivity analysis applying a two-species minimum interaction threshold. Lines represent predicted values from the location-scale GLMM. The analysis included networks with at least five honeybee individuals and wild pollinator species with at least five individuals per sampling event."}

library(ggeffects)
library(ggplot2)
library(patchwork)
library(glmmTMB)

# Load models
fit_SES_pdi <- readRDS("Data/fit_SES_pdi_over1.rds")
fit_SES_nd  <- readRDS("Data/fit_SES_nd_over1.rds")
fit_SES_mh  <- readRDS("Data/fit_SES_mh_over1.rds")

# Base theme
base_theme <- theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.key            = element_rect(fill = "white", colour = NA),
    legend.background     = element_blank(),
    legend.box.background = element_blank()
  )

my_cols <-  c(
  "#000000",  # black
  "#E69F00",  # orange
  #"#56B4E9",  # sky blue
  "#009E73",  # bluish green
  "#F0E442",  # yellow
  "#0072B2",  # blue
#  "#D55E00",  # vermillion
  "#CC79A7"   # reddish purple
)


# Marginal effects
gp_mh  <- ggpredict(fit_SES_mh,  terms = c("Dominance[all]", "Group"), type = "fixed")
gp_pdi <- ggpredict(fit_SES_pdi, terms = c("Dominance[all]", "Group"), type = "fixed")
gp_nd  <- ggpredict(fit_SES_nd,  terms = c("Dominance[all]", "Group"), type = "fixed")

# Plots with different linetypes per group
mh <- plot(gp_mh) +
  aes(linetype = group) +
    scale_color_manual(values = my_cols) +  
  labs(
    x = "",
    y = "SES Morisita-Horn",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",  # don't show CI fill as a separate legend
    colour   = guide_legend(override.aes = list(fill = NA)),   # <-- remove grey box
    linetype = guide_legend(override.aes = list(fill = NA))    # <-- same for linetype legend
  ) +
  ggtitle("")

pdi <- plot(gp_pdi) +
  aes(linetype = group) +
  scale_color_manual(values = my_cols) +  
  labs(
    x = "Honeybee dominance (ln-ratio)",
    y = "SES PDI",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",
    colour   = guide_legend(override.aes = list(fill = NA)),
    linetype = guide_legend(override.aes = list(fill = NA))
  ) +
  ggtitle("")

nd <- plot(gp_nd) +
  aes(linetype = group) +
  scale_color_manual(values = my_cols) +  
  labs(
    x = "",
    y = "SES normalised degree",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",
    colour   = guide_legend(override.aes = list(fill = NA)),
    linetype = guide_legend(override.aes = list(fill = NA))
  ) +
  ggtitle("")

# Combine and apply theme
p <- (mh | pdi | nd) +
  plot_layout(ncol = 3, nrow = 1, guides = "collect") &
  base_theme &
  theme(
    legend.position = "right",
    base_size = 7.5,
    panel.spacing = unit(0.1, "lines"),
    plot.margin = unit(c(0, 0, 0, 0), "pt"),
    plot.tag = element_text(face = "bold", vjust = 0.35),
    legend.key.width = unit(1.4, "cm")
  )

p + plot_annotation(tag_levels = "A", tag_suffix = ")")

```

\newpage

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%", fig.height = 2.5, fig.width = 8,  fig.cap= "\\textbf{Figure S3}. Effect of honeybee dominance on the standardised effect size (SES) of A) resource overlap (Morisita-Horn), B) PDI, and C) normalised degree, shown for the sensitivity analysis applying a nine-species minimum interaction threshold. Lines represent predicted values from the location-scale GLMM. The analysis included networks with at least five honeybee individuals and wild pollinator species with at least five individuals per sampling event."}
library(ggeffects)
library(ggplot2)
library(patchwork)
library(glmmTMB)
library(ggeffects)
library(ggplot2)
library(patchwork)
library(glmmTMB)

# Load models
fit_SES_pdi <- readRDS("Data/fit_SES_pdi_over9.rds")
fit_SES_nd  <- readRDS("Data/fit_SES_nd_over9.rds")
fit_SES_mh  <- readRDS("Data/fit_SES_mh_over9.rds")

# Base theme
base_theme <- theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.key            = element_rect(fill = "white", colour = NA),
    legend.background     = element_blank(),
    legend.box.background = element_blank()
  )

# Marginal effects
gp_mh  <- ggpredict(fit_SES_mh,  terms = c("Dominance[all]", "Group"), type = "fixed")
gp_pdi <- ggpredict(fit_SES_pdi, terms = c("Dominance[all]", "Group"), type = "fixed")
gp_nd  <- ggpredict(fit_SES_nd,  terms = c("Dominance[all]", "Group"), type = "fixed")

# Plots with different linetypes per group
mh <- plot(gp_mh) +
  aes(linetype = group) +
  scale_color_manual(values = my_cols) +  
  labs(
    x = "",
    y = "SES Morisita-Horn",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",  # don't show CI fill as a separate legend
    colour   = guide_legend(override.aes = list(fill = NA)),   # <-- remove grey box
    linetype = guide_legend(override.aes = list(fill = NA))    # <-- same for linetype legend
  ) +
  ggtitle("")

pdi <- plot(gp_pdi) +
  aes(linetype = group) +
  scale_color_manual(values = my_cols) +  
  labs(
    x = "Honeybee dominance (ln-ratio)",
    y = "SES PDI",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",
    colour   = guide_legend(override.aes = list(fill = NA)),
    linetype = guide_legend(override.aes = list(fill = NA))
  ) +
  ggtitle("")

nd <- plot(gp_nd) +
  aes(linetype = group) +
  scale_color_manual(values = my_cols) +  
  labs(
    x = "",
    y = "SES normalised degree",
    color    = "Group",
    linetype = "Group"
  ) +
  guides(
    fill     = "none",
    colour   = guide_legend(override.aes = list(fill = NA)),
    linetype = guide_legend(override.aes = list(fill = NA))
  ) +
  ggtitle("")

# Combine and apply theme
p <- (mh | pdi | nd) +
  plot_layout(ncol = 3, nrow = 1, guides = "collect") &
  base_theme &
  theme(
    legend.position = "right",
    base_size = 7.5,
    panel.spacing = unit(0.1, "lines"),
    plot.margin = unit(c(0, 0, 0, 0), "pt"),
    plot.tag = element_text(face = "bold", vjust = 0.35),
    legend.key.width = unit(1.4, "cm")
  )

p + plot_annotation(tag_levels = "A", tag_suffix = ")")
```

